Learning Network Size while Training with ShrinkNets

### 2018-02-09 paper/ICML/appendix.tex
Resolve even more references

\end{proof}
\subsection{Additional Experiment}
As we showed, Group sparsity share similarities with our method, and we claim
that ShrinkNets are a relaxation of group sparsity.  In this experiment we want
to compare the two aproaches.  We decided to focus on multi-target linear
regression because in the single target case, groups in the Group Sparsity
problem would have a size of one ($\bm{A}$ would be a vector in this case).

>>>>

\end{proof}
\subsection{Multi-Target Linear and Multi-Class Logistic regressions experiments}
As we showed, Group sparsity share similarities with our method, and we claim
that ShrinkNets are a relaxation of group sparsity.  In this experiment we want
to compare the two aproaches.  We decided to focus on multi-target linear
regression because in the single target case, groups in the Group Sparsity
problem would have a size of one ($\bm{A}$ would be a vector in this case).

### 2018-02-09 paper/ICML/evaluation.tex
Resolve even more references

Previous attempt at solving this problem focus on simple datasets and simple
architectures (limited to Fully connected). We believe that self-resizing
network really matter for problems where the architecture you need to solve them
involve so many layers that it is impractical to explore the space of possible
sizes. We considered ImageNet but the time required to train models is so long
that it would have not been possible to train enough and get statistically
significant results. This is why we picked \texttt{CIFAR10} \cite{Krizhevsky2009}. It is an
image classification dataset containing $60000$ color images $(3 \times 32
\times 32)$, belonging to $10$ different classes.

To solve this task we use the \texttt{VGG16} model \cite{Srivastava2014}. It is constituted of
alternating convolutional layers and \textit{MaxPool} layers interleaved by
\textit{BatchNorm} [ref] and \textit{ReLU} [ref] layers. The two last layers are
Fully connected layers separated by just a \textit{ReLU} activation function.

@@

ShrinkNets assume that the starting size of the network is an upper bound on the
optimal size. We though that picking two times the recommended size for each
layer (that was designed for ImageNet[ref]), is a generous upper bound. For the
classification layers we use $5000$ neurons as an upper bound where the ImageNet
version uses $4096$ \gl{This is on the top of my head, need to be double checked}.

We assume no prior knowledge on the optimal batch size, learning rate,
$\lambda$ or weight decay ($\lambda_2$). This is why, we randomly sample them
from a range of reasonable values (\gl{should we make them explicit ?}).
Training is done using our library, based on \texttt{PyTorch} \cite{paszke2017automatic}, that
support dynamically resized layers. We train the network using gradient descent
and \textit{Adam} optimizer \cite{DBLP:journals/corr/KingmaB14}. We start with
the learning rate sampled randomly and every $5$ epochs of non improvement in
validation accuracy we divide the learning rate by $10$. We stop training after
$400$ epochs or when the learning rate is under $10^{-7}$ whichever comes
first. \gl{Should I also give the details about the removal strategy we used
  and the $\gamma$ and threshold I used ? because this is clearly getting
  boring} For each of the models we trained, we pick the epoch with the best
validation accuracy and report the corresponding testing accuracy. Because of
the nature of our method, it can happen that for network that are aggressively
compressed, the best validation accuracy is obtained early in the training,
before the size has converged. To be sure that accuracy measured corresponds to
the final shape and not the starting shape, we only consider the second half of
the training when picking the best epoch. For each model, we also measure the
total size, in number of floating point parameters, excluding the
\textit{Switch Layers} because as we saw in section \gl{REFERENCE\_ME}, we can
get rid of them when training is done.

@@

The only difference in terms of setup are minimal. We use the same architecture
as \cite{Scardapane2017}, i.e. three layers Fully Connected neural network with  no
\textit{Dropout} \cite{Srivastava2014} nor \textit{BatchNorm} \cite{DBLP:journals/corr/IoffeS15}. \gl{Should we say here that we
  don't expect Dropout to work here ? I could write an entire paragraph about it
  if needed}. In this case for the \textit{Static Networks}, we sample
independently sizes of the three different layers to explore all possible
architectures.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{neuron_removal}
\vspace*{-5mm}
\caption{\label{neuron_removal_figure}Effect of dynamic neuron removal for different $\gamma$. First column is the difference in the final loss in function of the removal factor. We plot theoretical baseline as a reference. Right column is a proxy of the total cost for training the model (i.e. the sum of input neurons at each epoch). Each row is a dataset/$\lambda$ combination. From top to bottom we have: \texttt{scm1d}/$0.1$, \texttt{oes97}/$0.1$}

\end{center}
\vspace*{-4mm}
\end{figure}

@@

We sucessfully demonstrated that the architecture \textit{Shrinknets} converge
to make sense both in term of size and performance. And for a given accuracy the
size needed is significantly smaller than when we use the classic heursitc we
commonly use to size convolutional neural networks. It might be an indication
that this strategy is suboptimal and might need to be adjusted. During our
experimentations on simpler datasets like \texttt{MNIST} \cite{Lecun1998} and
\texttt{FashionMNIST} \cite{Xiao2017} we observed that even if they have the same number of
classes, input features, and output distribution. For a fixed $\lambda$
\textit{ShrinkNets} converged to considerably bigger networks in the case of
\texttt{FashionMNIST}. This evidence shows that final architecture not only
depends on the output distribution or shape of the data but actually reflects
the dataset (we know that FASHIONMNIST is harder then MNIST)

@@

\subsection{Multi-Target Linear and Multi-Class Logistic regressions [To be
  moved in the theoretical part]}

\subsection{Neuron Removal strategies [TO be removed]}

In our previous experiment, we showed that the ShrinkNet loss was
reasonable in practice, we are now interested in the impact on
early pruning. The method we suggest for early prunning uses a
parameter $\gamma$ that control the aggressiveness of neuron removal
so we will try to evaluate its impact on the final loss achieved by
the model and the cost required to train the model. Our cost model
is simple and hardware independant, we sum the number of input neurons
at each epoch. In theory the cost in time should be asymptotically
linear to this metric. To have a baseline we also train the same model
but without neuron removal. Keep in mind that this is just in order to
have some reference. Indeed, if we were to remove the neurons with small
weights it would deteriorate the loss (and picking the threshould would
be completely arbitrary). Therefore the baseline is evaluated with all neurons.
One could consider it as a "theoretical lower bound" of the best achievable loss.

We picked multiple combinations of dataset and regularization parameters ($\lambda$) and for each we fit with different aggressiveness parameters ($\gamma$). We measure the loss after convergence and the total cost and report the result in \cref{neuron_removal_figure}. In order to reduce the noise in the result, each experiment was performed $30$ times and we display the range arround $\pm$ $1$ standard deviation.

>>>>

Previous attempt at solving this problem focus on simple datasets and simple
architectures (limited to Fully connected). We believe that self-resizing
network really matter for problems where the architecture you need to solve them
involve so many layers that it is impractical to explore the space of possible
sizes. We considered ImageNet \cite{ILSVRC15} but the time required to train models is so long
that it would have not been possible to train enough and get statistically
significant results. This is why we picked \texttt{CIFAR10} \cite{Krizhevsky2009}. It is an
image classification dataset containing $60000$ color images $(3 \times 32
\times 32)$, belonging to $10$ different classes.

To solve this task we use the \texttt{VGG16} model \cite{Srivastava2014}. It is
constituted of alternating convolutional layers and \textit{MaxPool} layers
interleaved by \textit{BatchNorm} \cite{DBLP:journals/corr/IoffeS15} and
\textit{ReLU} \cite{Nair2010}  layers. The two last layers are Fully connected
layers separated by just a \textit{ReLU} activation function.

@@

ShrinkNets assume that the starting size of the network is an upper bound on the
optimal size. We though that picking two times the recommended size for each
layer (that was designed for ImageNet), is a generous upper bound. For the
classification layers we use $5000$ neurons as an upper bound where the ImageNet
version uses $4096$ \gl{This is on the top of my head, need to be double checked}.

We assume no prior knowledge on the optimal batch size, learning rate,
$\lambda$ or weight decay ($\lambda_2$). This is why, we randomly sample them
from a range of reasonable values (\gl{should we make them explicit ?}).
Training is done using our library, based on \texttt{PyTorch} \cite{paszke2017automatic}, that
support dynamically resized layers. We train the network using gradient descent
and \textit{Adam} optimizer \cite{DBLP:journals/corr/KingmaB14}. We start with
the learning rate sampled randomly and every $5$ epochs of non improvement in
validation accuracy we divide the learning rate by $10$. We stop training after
$400$ epochs or when the learning rate is under $10^{-7}$ whichever comes
first. \gl{Should I also give the details about the removal strategy we used
  and the $\gamma$ and threshold I used ? because this is clearly getting
  boring} For each of the models we trained, we pick the epoch with the best
validation accuracy and report the corresponding testing accuracy. Because of
the nature of our method, it can happen that for network that are aggressively
compressed, the best validation accuracy is obtained early in the training,
before the size has converged. To be sure that accuracy measured corresponds to
the final shape and not the starting shape, we only consider the second half of
the training when picking the best epoch. For each model, we also measure the
total size, in number of floating point parameters, excluding the
\textit{Switch Layers} because as we saw in \cref{neuron_killing}, we can
get rid of them when training is done.

@@

The only difference in terms of setup are minimal. We use the same architecture
as \cite{Scardapane2017}, i.e. three layers Fully Connected neural network with  no
\textit{Dropout} \cite{Srivastava2014} nor \textit{BatchNorm}. \gl{Should we say here that we
  don't expect Dropout to work here ? I could write an entire paragraph about it
  if needed}. In this case for the \textit{Static Networks}, we sample
independently sizes of the three different layers to explore all possible
architectures.

@@

% \begin{figure}
% \begin{center}
% \includegraphics[width=\columnwidth]{neuron_removal}
% \vspace*{-5mm}
% \caption{\label{neuron_removal_figure}Effect of dynamic neuron removal for different $\gamma$. First column is the difference in the final loss in function of the removal factor. We plot theoretical baseline as a reference. Right column is a proxy of the total cost for training the model (i.e. the sum of input neurons at each epoch). Each row is a dataset/$\lambda$ combination. From top to bottom we have: \texttt{scm1d}/$0.1$, \texttt{oes97}/$0.1$}
% 
% \end{center}
% \vspace*{-4mm}
% \end{figure}

@@

We sucessfully demonstrated that the architecture \textit{Shrinknets} converge
to make sense both in term of size and performance. And for a given accuracy
the size needed is significantly smaller than when we use the classic heursitc
we commonly use to size convolutional neural networks. It might be an
indication that this strategy is suboptimal and might need to be adjusted.
During our experimentations on simpler datasets like \texttt{MNIST}
\cite{Lecun1998} and \texttt{FashionMNIST} \cite{Xiao2017} we observed that
even if they have the same number of classes, input features, and output
distribution. For a fixed $\lambda$ \textit{ShrinkNets} converged to
considerably bigger networks in the case of \texttt{FashionMNIST}. This
evidence shows that final architecture not only depends on the output
distribution or shape of the data but actually reflects the dataset, indeed we
know that \texttt{MNIST} is a much easier problem than \texttt{FashionMNIST}.

@@

% \subsection{Neuron Removal strategies [TO be removed]}
% 
% In our previous experiment, we showed that the ShrinkNet loss was
% reasonable in practice, we are now interested in the impact on
% early pruning. The method we suggest for early prunning uses a
% parameter $\gamma$ that control the aggressiveness of neuron removal
% so we will try to evaluate its impact on the final loss achieved by
% the model and the cost required to train the model. Our cost model
% is simple and hardware independant, we sum the number of input neurons
% at each epoch. In theory the cost in time should be asymptotically
% linear to this metric. To have a baseline we also train the same model
% but without neuron removal. Keep in mind that this is just in order to
% have some reference. Indeed, if we were to remove the neurons with small
% weights it would deteriorate the loss (and picking the threshould would
% be completely arbitrary). Therefore the baseline is evaluated with all neurons.
% One could consider it as a "theoretical lower bound" of the best achievable loss.
% 
% 
% We picked multiple combinations of dataset and regularization parameters ($\lambda$) and for each we fit with different aggressiveness parameters ($\gamma$). We measure the loss after convergence and the total cost and report the result in \cref{neuron_removal_figure}. In order to reduce the noise in the result, each experiment was performed $30$ times and we display the range arround $\pm$ $1$ standard deviation.

### 2018-02-09 paper/ICML/impl.tex
Resolve even more references

\subsubsection{Neuron killing}
Once Switch layers are placed in a network and
initialized (sampled from the $\mathcal{N}(0, 1)$ distribution),

\textbf{Threshold strategy}:  We kill neurons based on a threshold (in absolute
value), this is the method used by deep compression. It is bad because it is
scale dependant and is not robust to noise in the gradients (explain that)
\\ \textbf{Sign change strategy}:  We look at when the sign change. This is also
sensitive to noise but has the advantage
\textbf{Sign variance strategy}: We measure the variance of the sign ($-1, 1$) of each component in the \textit{Switch Layer}, and consider it dead when it goes over a predefined threshold. \gl{Maybe we should add some intuition on why it makes sense}. \gl{We should probably explain the impact of $\gamma$ and report numbers from the expertiment that show that it does not hurt too much to kill neurons, we have the data in this experiment in the evaluation that we probably want to remove later.}

@@

We have implemented \textbf{Switch Layers} and the associated training procedure
as a library in pytorch~\cite{github}.
The layer can be freely mixed with other popular layers such as convolutional
layers, batchnorm layers, fully connected layers etc.
There are two key challenges that must be overcome while implementing 
\textbf{Switch Layers}.
First, since ShrinkNets kill off a large fraction of neurons, we must remove the
overhead of carrying around dead neurons.
And second, we must update the various network layers and training machinery (e.g.
optimizer state) to reflect the deactivation of neurons.
We address these challenges by augmenting every layer with additional information
about neuron status and implementing a \emph{neural garbage collection} 
mechanism which prunes deactivated neurons on-the-fly and updates the state of 
the training machinery.
% We do so by augmenting every layer with additional information such as what
% neurons have been deactivated.
% The layer then updates its size accordingly and this change is then propagated
% to other parts of the training machinery including the optimizer which must
% know the sizes of parameters.

>>>>

\subsection{Neuron killing}
\label{neuron_killing}
Once Switch layers are placed in a network and
initialized (sampled from the $\mathcal{N}(0, 1)$ distribution),

\textbf{Threshold strategy}:  We kill neurons based on a threshold (in absolute
value), this is the method used by deep compression. It is bad because it is
scale dependant and is not robust to noise in the gradients (explain that) \\

\textbf{Sign change strategy}:  When using this strategy we consider a neuron
dead when it changes its sign. This strategy works well in practice but is
sensible to noise. If we sample a gradient that is not representative of the
dataset when performing gradient descent then we might wrongly kill a neuron. \\

\textbf{Sign variance strategy}: We measure the exponnential moving variance of
the sign ($-1, 1$) of each component in the \textit{Switch Layer}, and consider
it dead when it goes over a predefined threshold. Thus, this strategy introduce
two extra parameters, one that controls the behavior of the innertia of the
variance, and the other is the threshold.

According to our experiments on linear and logistic regressions the
two last strategies reduce the number of parameter update performed during
training by at least an order of magnitude with only minimal impact on the final
error obtained by the model.

@@

We have implemented \textbf{Switch Layers} and the associated training procedure
as a library in pytorch~\cite{paszke2017automatic}.
The layer can be freely mixed with other popular layers such as convolutional
layers, batchnorm layers, fully connected layers etc.
There are two key challenges that must be overcome while implementing 
\textbf{Switch Layers}.
First, since ShrinkNets kill off a large fraction of neurons, we must remove the
overhead of carrying around dead neurons.
And second, we must update the various network layers and training machinery (e.g.
optimizer state) to reflect the deactivation of neurons.
We address these challenges by augmenting every layer with additional information
about neuron status and implementing a \emph{neural garbage collection} 
mechanism which prunes deactivated neurons on-the-fly and updates the state of 
the training machinery.
% We do so by augmenting every layer with additional information such as what
% neurons have been deactivated.
% The layer then updates its size accordingly and this change is then propagated
% to other parts of the training machinery including the optimizer which must
% know the sizes of parameters.

### 2018-02-09 paper/ICML/introduction.tex
Resolve even more references

Of course, the problem of finding compressed or small networks is not new.
Existing techniques typically aim make a pre-trained neural network smaller
\cite{something1,something2} by taking one of two approaches:  either by
applying quantitazation \cite{quant} or code compilation~\cite{something}
techniques that can be applied blindly to any network, or they analyze the
structure of the network and systematically prune
connections~\cite{han2015deepcompression,brain-damage} or neurons~\cite{XXX}
based on some loss function.  Although these techniques can substantially
reduce model size, they have several drawbacks.  First, they often negatively
impact model quality.  Second, they can (surprisingly) negatively impact
inference time as they transform dense matrix operations into sparse ones,
which can be substantially slower to execute on modern hardware
\cite{something}, especially GPUs which do not efficiently support sparse
linear algebra.  Third, these techniques generally start by optimizing
particular architecture for prediction performance, and then, as a post-
processing step, apply compression  to generate a smaller model that meets the
resource constraints of the   deployment setting.  Because the network
architecture is essentially fixed during compression,   model architectures
that work better in small settings may be missed -- this is especially true in
large networks like many- layered CNNs,  where it is infeasible to try explore
even a small fraction of possible network configurations.

@@

In contrast, in this paper we present a new method to simultaneously optimize
both network size and model performance. The key idea is to learn the right
network size at the same time that we optimize for prediction performance. Our
approach, called {\it ShrinkNets}, starts with an {\it oversized} network, and
dynamically shrinks it by eliminating neurons during training time.  Our
approach has two main benefits.  First, we explore the architecture of models
that are both small and perform well, rather than starting with a high-
performance model and making it small.  This allows us to efficiently generate
a family of smaller and accurate models without an exhaustive and expensive
hyperparameter search over the number of neurons in each layer. Second, in
contrast to existing neural network compression techniques~\cite{brain-
damage,han2015deepcompression}, our approach results in models that are both
not only small, but where the  layers are dense, which means that inference
time is also improved, e.g., on GPUs.

@@

In summary, our contributions are as follows: 
\begin{compactenum}
\item We propose a novel technique based on dynamically switching on and off neurons, 
which allows us to optimize the network size as the network is trained. 
\item We show that our technique is a relaxation\mpv{?} of group sparsity and 
prove \mpv{fill in}. 
\item \srm{Some claim about model size vs performance}
\item We demonstrate the efficacy of our technique on both convolutional and fully-connected neural nets,
showing that 
ShrinkNets finds networks within +/-X\% of best hand-crafted accuracy in XX\% of
training time compared to existing hyperparameter optimization methods.
\item We also demonstrate that ShrinkNets can achieve this accuracy with only YY\% 
of neurons.
\item \srn{Something about dense networks and then benefit over optimal brain damage}
\item \srm{Some claim about inference time}
\item \srm{Some claim about compatibility with existing compression techniques?}
\end{compactenum}

>>>>

Of course, the problem of finding compressed or small networks is not new.
Existing techniques typically aim make a pre-trained neural network smaller
\cite{something1,something2} by taking one of two approaches:  either by
applying quantitazation \cite{quant} or code compilation~\cite{something}
techniques that can be applied blindly to any network, or they analyze the
structure of the network and systematically prune
connections~\cite{han2015deepcompression,Cun} or neurons~\cite{XXX}
based on some loss function.  Although these techniques can substantially
reduce model size, they have several drawbacks.  First, they often negatively
impact model quality.  Second, they can (surprisingly) negatively impact
inference time as they transform dense matrix operations into sparse ones,
which can be substantially slower to execute on modern hardware
\cite{something}, especially GPUs which do not efficiently support sparse
linear algebra.  Third, these techniques generally start by optimizing
particular architecture for prediction performance, and then, as a post-
processing step, apply compression  to generate a smaller model that meets the
resource constraints of the   deployment setting.  Because the network
architecture is essentially fixed during compression,   model architectures
that work better in small settings may be missed -- this is especially true in
large networks like many- layered CNNs,  where it is infeasible to try explore
even a small fraction of possible network configurations.

@@

In contrast, in this paper we present a new method to simultaneously optimize
both network size and model performance. The key idea is to learn the right
network size at the same time that we optimize for prediction performance. Our
approach, called {\it ShrinkNets}, starts with an {\it oversized} network, and
dynamically shrinks it by eliminating neurons during training time.  Our
approach has two main benefits.  First, we explore the architecture of models
that are both small and perform well, rather than starting with a high-
performance model and making it small.  This allows us to efficiently generate
a family of smaller and accurate models without an exhaustive and expensive
hyperparameter search over the number of neurons in each layer. Second, in
contrast to existing neural network compression techniques~\cite{Aghasi2016,han2015deepcompression}, our approach results in models that are both
not only small, but where the  layers are dense, which means that inference
time is also improved, e.g., on GPUs.

@@

In summary, our contributions are as follows: 
\begin{compactenum}
\item We propose a novel technique based on dynamically switching on and off neurons, 
which allows us to optimize the network size as the network is trained. 
\item We show that our technique is a relaxation\mpv{?} of group sparsity and 
prove \mpv{fill in}. 
\item \srm{Some claim about model size vs performance}
\item We demonstrate the efficacy of our technique on both convolutional and fully-connected neural nets,
showing that 
ShrinkNets finds networks within +/-X\% of best hand-crafted accuracy in XX\% of
training time compared to existing hyperparameter optimization methods.
\item We also demonstrate that ShrinkNets can achieve this accuracy with only YY\% 
of neurons.
\item \srm{Something about dense networks and then benefit over optimal brain damage}
\item \srm{Some claim about inference time}
\item \srm{Some claim about compatibility with existing compression techniques?}
\end{compactenum}

### 2018-02-09 paper/ICML/related_work.tex
Resolve even more references

\begin{itemize}
  \item Hyperparameter optimization: random, bayesian opt, bandit methods
  \item distillation techniques
  \item post-training compression techniques
  \item group sparsity, non-parametric neural networks
  \item training dynamics paper: first overfitting and then randomization?
\end{itemize}

@@

In contrast with hyperparameter tuning methods, some methods such as 
DeepCompression~\cite{han2015deepcompression} seek to compress the network to 
make inference more efficient.
This is accomplished by pruning connections and quantizing weights.
On similar lines, multiple techniques such as~\cite{romero2014fitnets, 
hinton2015distilling} have been proposed for distilling a network into a 
simpler network or a different model.
Unlike our technique which works during training, these techiques are used after
training and it would be interesting to apply them to ShrinkNets as well.

The techniques closed to our work are those based on group sparsity such as 
\mpv{Guillaume: fill in}.

>>>>

\begin{itemize}
  \item Hyperparameter optimization: random, bayesian opt, bandit methods
  \item distillation techniques
  \item post-training compression techniques
  \item group sparsity, non-parametric neural networks
  \item training dynamics paper: first overfitting and then randomization?, \gl{Here is the ref, if you can introduce it in the flow \cite{Shwartz-Ziv2017}}
\end{itemize}

@@

In contrast with hyperparameter tuning methods, some methods such as
DeepCompression~\cite{han2015deepcompression} seek to compress the network to
make inference more efficient.  This is accomplished by pruning connections and
quantizing weights.  On similar lines, multiple techniques such
as~\cite{romero2014fitnets, hinton2015distilling} have been proposed for
distilling a network into a simpler network or a different model.  Unlike our
technique which works during training, these techiques are used after training
and it would be interesting to apply them to ShrinkNets as well. \cite{Abadi2016b} share the common goal of removing entire blocks of parameter to maintain dense matrices, however their method only applies to convolutional layers.

The techniques closed to our work are those based on group sparsity such as
~\cite{Scardapane2017}, and those like ~\cite{Philipp} who grow and shrink
dynamically the size of the network during training.
